---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# setaria-predict

<!-- badges: start -->
<!-- badges: end -->

The goal of setaria-predict is to train a model using simulation output from ED2 to create an ED2 "emulator" of sorts.  Then, we can use this model to predict the growth of *Setaria* at new locations.

We will use runs from the sensitivity analysis as stand-ins for different *Setaria* genotypes and train models using site location and/or weather data as predictors, then predict aboveground biomass or NPP of *Setaria* at new sites.

## Reproducibility

This project uses [`renv`](https://rstudio.github.io/renv/articles/renv.html) for package management. When opening this repo as an RStudio Project for the first time, `renv` should automatically install itself and prompt you to run `renv::restore()` to install all package dependencies.

This project uses the [`targets` package](https://docs.ropensci.org/targets/) for workflow management. Run `targets::tar_make()` from the console to run the workflow and reproduce all results.

Data for this project was originally on a server (named Welsch) in many CSV files (one per simulation).  These are not committed in this repository, but the combined data is committed to `data/` and the script used to wrangle them is in `data_welsch/`.

**Targets workflow:**

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
cat(
  "```mermaid",
  targets::tar_mermaid(targets_only = FALSE),
  "```",
  sep = "\n"
)
#this should display correctly on GitHub, or code can be pasted into https://mermaid.live
```