---
title: "Exploratory Data Analysis"
format: html
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
library(skimr)
library(targets)
library(tidyverse)
tar_load(model_data)

```

Look at distribution of response variable (summer NPP)

```{r}
ggplot(model_data, aes(npp_summer)) +
  geom_histogram() +
  facet_wrap(~ecosystem)
```

-   Are negative values meaningful? (sounds like the answer is "it depends" [@roxburgh2005].

-   If not, should they be converted to zeros?

-   If they can be converted to zeros, should data be log-transformed (or modeled with a log-link)?

```{r}
colnames(model_data)
skim(model_data)
```

-   `mean_swe` is invariant at 0 and should be removed

-   `year` should probably be treated as categorical

```{r}
ggplot(model_data, aes(year, npp_summer)) +
  geom_point() +
  facet_grid(~ecosystem)
```

```{r}
ggplot(model_data, aes(mean_srad, mean_dayl)) + geom_point()
cor.test(model_data$mean_dayl, model_data$mean_srad)
```

Mean solar radiation is correlated with mean day length, but they might contain slightly different information. BioClim variables might be better.
