---
title: "Predictive Modeling of Setaria NPP"
author: "Eric R. Scott"
format: html
editor: visual
bibliography: references.bib
csl: ecology.csl
---

```{r}
# source(here::here(".Rprofile"))
library(targets)
library(tarchetypes)
library(sf)
library(tidyverse)
library(tidymodels)
library(patchwork)
library(ragg)
library(ggtext)
tar_load(c(
  rf_recipe,
  data_split,
  lm_cv,
  lm_pred_plot,
  rf_cv,
  rf_pred_plot
))
```

## Methods Overview

The goal was to generate predictive spatial models for four simulated phenotypes of\
*Setaria*: wildtype, high anthocyanin content, dwarf, and increased leaf temperature.
We used PEcAn to generate ED2 runs at 29 southeast US sites in 3 ecosystems at each site: mixed forest, pine forest, and prairie.
We utilized PEcAn's sensitivity analysis to simulate the 4 phenotypes.
The median run was used as wildtype, -1 SD for quantum efficiency represented high anthocyanin content, -1 SD for stomatal slope represented the increased leaf temperature phenotype, and +1 SD for fine root to leaf ratio represented the dwarf phenotype.
We used only the final year of the 10 year simulations as data, averaging the NPP over the summer months (June, July, and August) as a response variable.
We downloaded weather data for each site from Daymet using the `daymetr` package [@daymetr] and calculated 19 bioclimatic variables using the `dismo` package [@dismo].
The resulting dataset contained mean summer NPP, the site ID, the 19 bioclimatic variables, mean solar radiation, and mean vapor pressure deficit, ecosystem, and phenotype.
Data preparation, model training, model validation, and prediction were facilitated by the packages in the `tidymodels` ecosystem [@tidymodels].
The dataset was split 80/20 into a training and test set using the `rsample` package [@rsample] and prepared for modeling using the `recipes` package [@recipes] (steps detailed below).
We compared two candidate models, a random forest model (using the `ranger` engine), and a LASSO regression (using the `glmnet` engine) [@glmnet; @ranger; @tay2023].
We compared the predictive ability of the candidate models using 10-fold cross-validation RMSE and $R^2$ values.\

A grid of points every 0.5ยบ was generated within the bounds of the Southeastern US states that the ED2 simulations were sampled from using the `terra`, `stars`, and `sf` packages [@terra; @stars; @sf].

## Data Prep

```{r}
#| collapse: true
rf_recipe
```

All numeric predictors (the 19 bioclim variables, mean daylight, and mean vapor pressure deficit) are centered and scaled, dummy variables are created for phenotype, ecosystem, and their interaction, and PCA is performed on the bioclim variables to retain a number of components that explains 85% of variance.

## Model Validation

I fit two models using the same data recipe, a random forest model and a LASSO regression and performed 10-fold cross validation.
Random forest outperformed LASSO regression.

```{r}
bind_rows("random forest" = rf_cv, "LASSO regression" = lm_cv, .id = "model") |> 
  dplyr::select(model, .metric, mean) |> 
  pivot_wider(values_from = mean, names_from = .metric)

```

```{r}
#| label: fig-pred_diagnostic
(rf_pred_plot + labs(title = "Random forest")) /
  (lm_pred_plot + labs(title = "LASSO regression")) + plot_layout(guides = "collect")
```

## Spatial predictions

I generated a grid of points to predict summer average NPP after 10 years of growth at.

```{r}
#| label: fig-grid
#| fig-cap: "Grid of points 0.5ยบ apart used for predicting mean summer NPP"
tar_load(c(seus, grid_data))
ggplot() + 
  geom_sf(data = seus) +
  geom_point(data = grid_data, aes(x = lon, y = lat))
```

The resulting predicted values look are plotted below for each ecosystem and phenotype

```{r}
#| label: pred-npp-map
#| fig.cap: Predicted summer NPP for four phenotypes of *Setaria* in three simulated ecosystems. NPP values were predicted using a random forest model trained on ED2 simulations at 29 sites and 3 ecosystems per site.  NPP values are predicted on a 0.5ยบ grid and displayed on a log scale.
tar_read(pred_map)
```

We can also visualize these as deviations from wildtype

```{r}
#| label: pred_npp_dif
#| fig.cap: Percent difference in summer NPP compared to wildtype *Setaria* predicted in the South East US.
tar_read(pred_map_diff)
```

## References

::: {#refs}
:::
