---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(targets)
library(tarchetypes)
tar_load(c(
  rf_recipe,
  data_split,
  lm_cv,
  lm_pred_plot,
  rf_cv,
  rf_pred_plot
))
```

## Data Prep

```{r}
#| collapse: true
rf_recipe
```

All numeric predictors (the 19 bioclim variables, mean daylight, and mean vapor pressure deficit) are centered and scaled, dummy variables are created for genotype, ecosystem, and their interaction, and PCA is performed on the bioclim variables to retain a number of components that explains 85% of variance.

The data were split 80/20 into training and test set and 10-fold CV datasets were created from the training set.

```{r}
data_split
```


## Model Validation

I fit two models using the same data recipe, a random forest model and a LASSO regression and performed 10-fold cross validation. Random forest outperformed LASSO regression.

```{r}
bind_rows("random forest" = rf_cv, "LASSO regression" = lm_cv, .id = "model") |> 
  dplyr::select(model, .metric, mean) |> 
  pivot_wider(values_from = mean, names_from = .metric)

```

```{r}
library(patchwork)
(rf_pred_plot + labs(title = "Random forest")) /
  (lm_pred_plot + labs(title = "LASSO regression")) + plot_layout(guides = "collect")
```

